Closing the empirical cycle

### 2022-11-10 action_letter.Rmd
update 2022-11-10 11:47:07

---
title       : "Using machine learning to identify early predictors of adolescent emotion regulation development"
authors     : "masked"
journal     : "Journal of Research on Adolescence"
editor      : "Raval"
manuscript  : "2022-182"
class       : "draft"
bibliography: ["manuscript/veni.bib"]
output      : papaja::revision_letter_pdf
---

@@

Yours sincerely,
Author

@@

We completely agree with Reviewer 3 that there is already a wealth of prior research in this area.
In fact, there is so much work on this topic that it would be a tall order to review the literature on an individual study basis.
Given the large prior literature,
we rely primarily on prior review articles.
We currently cite 17 review articles in the introduction,
including a recent (2021) comprehensive systematic review of 6305 studies on factors associated with adolescent emotion regulation.

We cannot strive to recreate a comprehensive review here - 
and this also seems redundant, given that many of the cited review articles are very recent (i.e., from 2019-2021),
and seems overly ambitious, given the page limits and attention that must be devoted to explaining the novel methods,
and spurious, given that this is an exploratory study and thus has no hypotheses that must be supported with prior literature.
The present study builds upon prior reviews of the literature,
and those reviews are cited.

Of course we are willing to include specific omitted literature, but please note that Reviewer 3 does not mention which references they feel are missing.

`r get_revision(manuscript = man, "abundantlit")`

@@

As mentioned above, the vast majority of cited studies are reviews (narrative reviews, meta-analyses, and book chapters) or theoretical publications.
Although one would hope that such publications cover all available literature without selecting for country of origin,
some bias is likely present.
Of the few primary empirical studies we cited, all originated in the "Global North" (if one also considers Israel and Greece to be part of the Global North).
To address this comment, we have extracted information about the country of origin of the samples of all primary data sources cited.
This information is summarized in a new Supplemental Table,
which also indicates what references are *not* primary data sources but instead theoretical, review, or methodological articles.

@@

@@

To address this comment, we have incorporated the following in our analysis of the results in light of the bioecological model:

@@

By contrast, parent-related constructs refer to any variable pertaining to a parent but that is not a parenting practice.
The highest ranking parent-related construct is balanced relatedness (3),
which measures whether the adolescent perceives his or her parents to be tolerant of the adolescent's differing beliefs.
Balanced relatedness is an aspect of the emotional family climate [@morrisRoleFamilyContext2007],
not a parenting practice, because it does not relate to specific parental actions.
The specific parenting practices associated with instilling a sense of balanced relatedness are maternal (54) and paternal (55) support, which rank very low.

The observation that *parenting practices* are relatively less important than individual predictors (and certainly much less important than we would have thought a-priori) is thus strongly supported by the data;
the first such behaviors are (16) and (25).
This is consistent with the finding that some parent-related constructs indeed score highly.

@@

We thank the Editor for this excellent suggestion; we have addressed this comment by adding the following paragraph:

@@

```{r figviolin, results='asis', out.width="20%", fig.cap="Variable importances clustered by theme (A) and level of the bioecological model (B).", eval = FALSE}
knitr::include_graphics("violin.png")
```

@@

We respectfully believe that this comment might be based on a misunderstanding. There are three types of missingness: MNAR, where missingness depends on unobserved data; MCAR, where missingness depends on observed data; and MAR, where missingness is truly random. It is not possible to test for MNAR, and the present methods are valid for both MCAR and MAR - so there is no need to provide evidence for either of those.

Also note the unfeasibility of performing an MCAR test when the number of variables is higher than the number of participants: The raw data, prior to imputation and scale construction, is 495 (participants) x 750 (variables) big. I am not aware of any MCAR test that works with N < p data.

@@

1. We added the requested illustrative example, using neuroticism. We could use any scale, but as Neuroticism later turns out to be the most important predictor, it seems a suitable candidate.
2. We have attempted to clarify our description of the dimension reduction procedure
3. We have added further statistics that, hopefully, give the reader a better sense of whether the use of these principal components is defensible. Specifically, we no longer just provide the explained variance by each first component - but we also standardize this explained variance by the number of items in the scale to see how much more variance the first component explains than expected by chance. We see that, across all scales, the first component explains between 1.6 - 17.46 times more variance than expected by chance.
</span>

@@

The Reviewer is correct; homoscedastic residuals were assumed because otherwise the model would be "hopelessly complex". We would have more parameters for the residuals than for the rest of the model, and the SEM-forest analysis would thus be primarily driven to explain differences in residuals, instead of in growth curves.
Unfortunately there is no way to test this assumption. The SEM-forest method estimates tens, or hundreds of thousands of SEM models in its different nodes. It is not possible to test model assumptions in each of those nodes.

@@

`r get_revision(manuscript = man, "examplenonlinear")`

@@

\RC{I know ensuring the article remains inside page limits is always a battle, but I had a suggestion for one additional, hopefully simplistic analysis that might even further bridge the “gap” between these machine learning analyses and “traditional, frequentist” analyses. Perhaps the authors could take only the most important predictors, say the top 10 (or really whatever number they want), and run regression analyses that predict trajectory intercepts and slopes? Basically, I know that people like me really like explanations like “oh yeah, these top 10 machine learning predictors predict XX% of variance in emotion regulation scores over time”, and so I was trying to think of a way to generate those sorts of tangible, “traditional” effect sizes. Could the authors potentially provide such effect size estimates? I also understand if there simply isn’t time or space for this, but figured it was worth asking.}

@@

\RC{I appreciated the brave and nuanced discussion of how these analyses conform to, and also challenge, traditionally held norms in the field of emotion regulation. I especially appreciated the identification that parenting, substance use, and bullying might not be the risk factors we think they are.
• The ERRATA tool seems amazing; I couldn’t find it online. Could a link be provided directly in the text?}

@@

We apologize for the inconvenience. The manuscript only has 42 pages; pages 43-65 are copies of the online supplemental materials (which are supposed to be accessed via the broken links, which have now been fixed). For the sake of blind review, we also submitted copies of these tables through the submission portal, but it evidently cannot correctly present these materials. The tables will not be part of the paper in this form.

@@

```{r}
knitr::include_graphics("violin.png")
```

@@

We also believe that the Figure above helps clarify these results.

@@

We do believe that the quoted statement is correct and does not require further modification, however, as it does not relate to causal interpretation of the results, but to the affordances of the method: latent class analysis *cannot* answer the question "Why?" some adolescents follow similar causal trajectories,
whereas SEM-forests *can* answer this question, because cases are clustered based on predictor variables (of course with the caveat that causal interpretation is contingent on study design or theoretical justification).

@@

1. Computing univariate partial dependence profiles is extremely computationally expensive; bivariate partial dependence is prohibitively so.
2. There is no a-priori way to determine which bivariate interactions might be important in SEM-forests analysis (because there is no applicable analogy for Friedman's H statistic).
3. Consequently, one would have to examine all bivariate interactions by hand. This comes down to interpreting 84*84*3 = 21168 plots. This is not possible.

@@

Again, we thank the Reviewer for the suggestion;
in principle this is correct, but in practice, no method (currently?) exists to obtain accumulated local effects for SEM-forests.
The Reviewer is citing cutting-edge developments in the area of classic random forests; it will take a while before these developments spill over into the SEM-forests literature, if ever.

@@

@@

@@

We apologize for the inconvenience. These Supplemental Tables will not be part of the printed manuscript; we were required to attach hard copies of the online supplemental tables in the submission portal, and these were evidently not formatted correctly. They are just spreadsheets that can be opened in a web-browser or excel.

@@

We appreciate these suggestions and the caution about abstract length; we have cut the abstract by half and attempted to incorporate the Reviewer's suggestions. The Abstract now reads:

@@

We appreciate this comment. There are two issues here, which we will both address in turn.

The first issue is that the offending paragraph evidently failed to convey that we set out to identify predictors of adolescents' diverging developmental trajectories using *person-centered* methods.
This is certainly different from preceding work, which has nearly exclusively looked at whole sample average effects.
This paragraph thus doesn't argue that "few have examined predictors of emotion regulation in adolescence", as the Reviewer interpreted it, but instead, argues that "few have used person-centered methods to examined predictors of adolescents' diverging developmental trajectories of emotion regulation."
In the Revision, we have attempted to clarify this by more precisely describing the contribution of the present study, in several places:

`r get_revision(manuscript = man, "personcentered1")`

The second issue is whether our writing does sufficient justice to prior publications on predictors of emotion regulation.
We could not agree more with the Reviewer that there is a wealth of prior research in this area.
In fact, there is so much work on this topic that it would be a tall order to do it all justice.
At present, we cite a total of 17 review articles in the introduction,
including a recent (2021) comprehensive systematic review of 6305 studies on factors associated with adolescent emotion regulation.

Given the large prior literature,
we found it necessary to rely primarily on review articles to do justice to prior work.
We cannot strive to recreate a comprehensive review here - 
and this also seems redundant, given that many of the cited review articles are very recent (i.e., from 2019-2021),
and seems overly ambitious, given the page limits and attention that must be devoted to explaining the novel methods,
and spurious, given that this is an exploratory study and thus has no hypotheses that must be supported with prior literature.
The present study builds upon prior reviews of the literature,
and those reviews are cited.

Of course we are willing to include **specific** omitted literature, but it is not clear from the comment which specific references should be included.

@@

`r get_revision(manuscript = man, "abundantlit")`

@@

Comprehensive descriptive statistics about all available demographic characteristics are reported in Tables 1 and 2.
To address this comment, we now stress this in the Participants section:

@@

>  From 2006 to 2011, trained interviewers
conducted annual home visits to collect questionnaire data. All independent variables were
assessed in the first measurement wave.

@@

We thank the Reviewer for pointing out this unfortunate choice of words, which should have been flagged during proof reading.
Entropy is not a criterion for class enumeration, but for evaluating the quality of the chosen solution (see Jung \& Wickrama, 2008). In the Revision, we have attempted to clarify this as follows:

@@

In line with the Reviewer's expectation, we did in fact use the median and 16th and 84th percentile.
The confusion arises from our attempt to explain to the readership what these percentiles mean, by drawing a comparison to +/- 1SD.

@@

`r get_revision(manuscript = man, "highrankingparent", quote = F)`

>>>>

---
title       : "Using machine learning to identify early predictors of adolescent emotion regulation development"
authors     : "masked"
journal     : "Journal of Research on Adolescence"
editor      : "Raval"
manuscript  : "2022-182"
class       : "draft"
bibliography      : ["manuscript/r-references.bib", "manuscript/references.bib", "manuscript/Veni.bib"]
output      : papaja::revision_letter_pdf
---

@@

Yours sincerely,

the authors

@@

We completely agree with Reviewer 3 that there is already a wealth of prior research in this area.

`r get_revision("priorlit", letter, quote = F)`

@@

First, as mentioned above, the vast majority of cited studies are reviews (narrative reviews, meta-analyses, and book chapters) or theoretical publications.
Although one would hope that such publications cover all available literature without selecting for country of origin,
some bias is likely present.
Of the few primary empirical studies we cited, all originated in the "Global North" (if one also considers Israel and Greece to be part of the Global North).
To address this comment, we have extracted information about the country of origin of the samples of all primary data sources cited.
This information is summarized in a new Supplemental Table,
which also indicates what references are *not* primary data sources but instead theoretical, review, or methodological articles.

@@

<!-- Second, this is not a hypothesis-testing study. It doesn't matter whether findings from previous samples generalize to the present one, because we do not deduce hypotheses from previous work. All that matters is whether our findings can be generalized to other samples. -->

@@

To address this comment, we have incorporated the following into the Introduction:

`r get_revision(manuscript = man, "velez1")`

And the following in the Discussion:

@@

By contrast, parent-related constructs refer to any variable pertaining to a parent but that is not a parenting practice.
The highest ranking parent-related construct is balanced relatedness (3),
which measures whether the adolescent perceives his or her parents to be tolerant of the adolescent's differing beliefs.
Balanced relatedness is an aspect of the emotional family climate [@morrisRoleFamilyContext2007],
not a parenting practice, because it does not relate to specific parental actions.
The parenting practices associated with instilling a sense of balanced relatedness are maternal (54) and paternal (55) support, which rank very low.

The observation that parenting practices are relatively less important than individual predictors (and certainly much less important than we would have thought a-priori) is thus strongly supported by the data;
the first such behaviors are (16) and (25).
This conclusion is not contradicted by the finding that some parent-related constructs score highly.

@@

We thank the Editor for this excellent suggestion; we have addressed this comment by adding the following to the Introduction:

`r get_revision(manuscript = man, "velez1")`

And the following to the Discussion:

@@

@@

We respectfully believe that this comment might be based on a misunderstanding. There are three types of missingness: MCAR, where missingness depends on observed data and those observed data can be used to impute; MAR, where missingness is truly random and it doesn't matter if observed data is used to impute; and MNAR, where missingness depends on unobserved data and can therefore neither be detected nor correctly imputed.

No test exists to exclude MNAR. For example, Little's MCAR test and non-parametric alternatives like Jamshidian & Jalal's non-parametric test can provide *evidence confirming MAR*, but the presence or absence of MAR does not exclude MNAR.

The present methods are valid for both MCAR and MAR - so there is no need to provide evidence for either of those.

Also note that no MCAR test exists for cases where the number of variables is higher than the number of participants (N < p): In this case, the raw data, prior to imputation and scale construction, is 495 (participants) x 750 (variables) big, so N < p.

@@

1. We added the requested illustrative example, using neuroticism. We could use any scale, but as Neuroticism later turns out to be the most important predictor, it seems a suitable candidate.
2. We have clarified our description of the dimension reduction procedure
3. We have added further statistics that, hopefully, give the reader a better sense of whether the use of these principal components is defensible. Specifically, we no longer just provide the explained variance by each first component - but we also standardize this explained variance by the number of items in the scale to see how much more variance the first component explains than expected by chance. We see that, across all scales, the first component explains between 1.6 - 17.46 times more variance than expected by chance.
</span>

@@

The Reviewer is correct; homoscedastic residuals were assumed because otherwise the model would be "hopelessly complex". We would have more parameters for the residuals than for the rest of the model, and the SEM-forest analysis would thus be primarily driven to explain differences in residuals, instead of in growth curves.
Unfortunately there is no way to test this assumption. The SEM-forest method estimates tens to hundreds of thousands of SEM models in across all nodes in all trees. It is not possible to test model assumptions in each of those nodes.

@@

`r get_revision(manuscript = man, "nonlinearexample")`

@@

\RC{I know ensuring the article remains inside page limits is always a battle, but I had a suggestion for one additional, hopefully simplistic analysis that might even further bridge the “gap” between these machine learning analyses and “traditional, frequentist” analyses. Perhaps the authors could take only the most important predictors, say the top 10 (or really whatever number they want), and run regression analyses that predict trajectory intercepts and slopes? Basically, I know that people like me really like explanations like “oh yeah, these top 10 machine learning predictors predict XX\% of variance in emotion regulation scores over time”, and so I was trying to think of a way to generate those sorts of tangible, “traditional” effect sizes. Could the authors potentially provide such effect size estimates? I also understand if there simply isn’t time or space for this, but figured it was worth asking.}

@@

\RC{I appreciated the brave and nuanced discussion of how these analyses conform to, and also challenge, traditionally held norms in the field of emotion regulation. I especially appreciated the identification that parenting, substance use, and bullying might not be the risk factors we think they are.}

We thank the Reviewer for these kind words. We share the Reviewer's enthusiasm for the potential change in perspective implied by these findings - although we understand the importance of tempering that enthusiasm as this is just a single preliminary study, with considerable limitations.

\RC{The ERRATA tool seems amazing; I couldn’t find it online. Could a link be provided directly in the text?}

@@

We apologize for the inconvenience. The manuscript only has 42 pages; pages 43-65 are copies of the online supplemental materials (the same Tables that were supposed to be accessed via the broken links, which are now fixed). For the sake of blind review, we also submitted copies of these tables through the submission portal, but it evidently cannot correctly present these materials. The tables will not be part of the paper in this form.

@@

@@

First, we really want to stress that we are not biased against parenting; much of our prior work has focused on the effect of specific parenting practices on child development.
This notwithstanding, we were legitimately surprised that these same specific parenting behaviors ranked surprisingly low in machine learning analysis - although we understand and acknowledge that this is just a single preliminary study, with considerable limitations.

Second, with regard to the Reviewer's statement that "the third-most important predictor being a positive
parenting behavior". Balanced relatedness is not a parenting practice (i.e., a specific action parents engage in as part of child rearing). Rather, it is an aspect of the (adolescent-perceived) emotional family climate. `r get_revision(manuscript = letter, "parents", quote = F)`

We also believe that Figure 3 helps clarify these results.

@@

We do believe that the quoted statement is correct and does not require further modification, however, as it does not relate to causal interpretation of the results, but to the affordances of the method: latent class analysis *cannot* answer the question "Why?" some adolescents follow similar causal trajectories,
whereas SEM-forests *can* answer this question, because cases are clustered based on predictor variables.
As is the case for all statistical analyses, this comes with the caveat that causal interpretation is contingent on study design or theoretical justification.

@@

1. Computing univariate partial dependence profiles is extremely computationally expensive; bivariate partial dependence is prohibitively so.
2. There is no a-priori way to determine which bivariate interactions might be important in SEM-forests analysis (because there is no applicable analogy for Friedman's H statistic).
3. Consequently, one would have to examine all bivariate interactions by hand. This comes down to interpreting 84\*84\*3 = 21168 plots. This is not possible.

@@

Again, we thank the Reviewer for the suggestion;
in principle this is correct, but in practice, no method (currently?) exists to obtain accumulated local effects for SEM-forests.
The Reviewer is citing cutting-edge developments in the area of classic random forests, which are not yet widely adopted or implemented for different algorithms;
it will take a while before these developments spill over into the SEM-forests literature.

@@

We have attempted to address this comment by explaining our decision to assume homoscedasticity:

@@

And also, by addressing it in the Limitations section:

@@

We apologize for the inconvenience. These Supplemental Tables will not be part of the printed manuscript; we were required to attach hard copies of the online supplemental tables in the submission portal, and these were evidently not formatted correctly. They are just spreadsheets that can be opened in a web-browser or excel. The links to these Supplemental Tables in the manuscript now work.

@@

We appreciate these suggestions and the caution about abstract length; we have cut the abstract by half and incorporated the Reviewer's suggestions. The Abstract now reads:

@@

First, we agree that this choice of words was unfortunate.
We did not mean to imply that such research has not yet been conducted.
The purpose of this sentence was to emphasize that it is *important* to examine risk factors and lead into the following 3-page section which addresses what is already known.
To clarify this, we have rewritten this sentence and the next as follows:

> `r get_revision("nextstep", man, quote = FALSE)` [...] `r get_revision("fragmented", man, quote = FALSE)`

A second issue is whether our writing sufficiently acknowledges prior research on predictors of emotion regulation.
We could not agree more with the Reviewer that there is a wealth of prior research in this area.
<span id = "priorlit">In fact, a recent (2021) systematic review identified 6305 studies on factors associated with adolescent emotion regulation.
Given this exceedingly large volume of prior literature,
we found it necessary to rely primarily on review articles and theoretical articles to do justice to prior work.
We currently cite 17 review articles and 12 theoretical pieces,
none of which were restricted to any one lab.
We thus respectfully disagree that we only address studies from a "small number of labs".

We cannot strive to recreate a comprehensive review here - 
and this also seems redundant, given that many of the cited review articles are very recent (i.e., from 2019-2021),
and overly ambitious, given the page limits and attention that must be devoted to explaining the novel methods,
and spurious, given that this is an exploratory study and thus has no hypotheses that must be supported with prior literature.

This notwithstanding, we are willing to include **specific** omitted literature, but it is not clear from the Reviewer's comment which specific references should be included.

@@

`r get_revision(manuscript = man, "abundantlit")`
</span>

@@

Comprehensive descriptive statistics about all available demographic characteristics were already reported in Tables 1 and 2.
However, we see that this was not made sufficiently clear.
To address this comment, in the revision, we now state the following:

@@

>  From 2006 to 2011, trained interviewers
conducted annual home visits to collect questionnaire data.

@@

We thank the Reviewer for pointing out this unfortunate choice of words, which should have been flagged during proof reading.
In fact, entropy is not a criterion for class enumeration, but for evaluating the quality of the chosen solution (see Jung \& Wickrama, 2008). In the Revision, we have attempted to clarify this as follows:

@@

In line with the Reviewer's expectation, we did use the median and 16th and 84th percentile.
The confusion arises from our attempt to explain to the readership what these percentiles mean, by pointing out that these percentiles correspond to +/- 1SD of a standard normal distribution.

@@

`r get_revision(manuscript = man, "highrankingparent")`
